
## Pratical Machines Learning Project-Assignment Writeup

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

### Data Cleaning
Before starting the project, load machine learning and plot packages that will be used in this project. 
```{r}
>library(caret)
>library(rpart)
>library(rattle)
>library(rpart.plot)
>library(randomForest)
```
First, we should load both training data and testing data into R. Remove error string and empty cells and replace with NA
```{r}
train_source="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_source="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_document=read.csv(url(train_source), na.strings=c("NA","#DIV/0!",""))
testing_document=read.csv(url(test_source), na.strings=c("NA","#DIV/0!",""))
View(training_document)
[1] 19622   160
View(testing_document)
[1]  20 160
```

### Cross Validation
After loading the data, it shows that training set has 19622 rows of data and prediction set has 20 rows of data with 160 variables. In order to cross validate data, we use "createDataPartition" function to split the training data into two seperate sub datasets, one is training set with 60% of data and another one is testing set with 40% of data.

```{r}
intrain=createDataPartition(training_document $classe, p=0.6, list=FALSE)
training = training_document [intrain, ]
testing= training_document [-intrain, ]
dim(training); dim(testing)
[1] 11776   160
[1] 7846  160
```
We will use training set to find the best model and use testing set to verify our findings, and use the model we selected according to training and testing set to predict the "classe" in the prediction set. 

Some of the variables have no variability, use nearZeroVar function in caret package to identify the variables with very little variability and would not likely to become predictors. And then remove these variables. 

```{r}
> badcols=nearZeroVar(training)
> training1=training[,-badcols]
> testing1=testing[,-badcols]
> nsvvariable=names(training1)
> nsvvariable
> dim(training1)
> dim(testing1)
[1] 11776   132
[1] 7846  132


> finaltesting1=testing_document[,-badcols]
> dim(finaltesting1)
[1]  20 132
```
After removing these new zero variance variables, we narrow it down to 132 variables for training, testing and prediction set. But some of the variables have no single value in it, so next step is to remove these variables with missing values.

```{r}
> missing=which(colSums(is.na(training1))>0)
> training2=training1[,-c(missing)]
> dim(training2)
 [1] 11776    59
 
> missing=which(colSums(is.na(testing1))>0)
> testing2=testing1[,-c(missing)]
> dim(testing2)
[1] 7846   59

> missing=which(colSums(is.na(finaltesting1))>0)
> finaltesting2=finaltesting1[,-c(missing)]
> dim(finaltesting2)
[1] 20 59
```
Now, we only have 59 variables in these three sets. But the first column is a 

```{r}
> training2=training2[c(-1)]
> dim(training2)
[1] 11776    58
> testing2=testing2[c(-1)]
> dim(testing2)
[1] 7846   58
> finaltesting=finaltesting2[c(-1, -58)]
> dim(finaltesting)
[1] 20 57
```

### Model selection and evaluation
After rerange the data into proporiate format, we can use training set to fit the model. The first model we choose is the random forest. 

```{r}
> modelfit1=randomForest(classe ~ ., data=training2)
```
We can use this modelfit to predict the testing dataset to see accuracy of the prediction.
```{r}
> predictions1=predict(modelfit1, testing2, type = "class")
> confusionMatrix(predictions1, testing2$classe)

Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 2232    0    0    0    0
         B    0 1518    3    0    0
         C    0    0 1363    6    0
         D    0    0    2 1280    2
         E    0    0    0    0 1440

Overall Statistics
                                          
               Accuracy : 0.9983          
                 95% CI : (0.9972, 0.9991)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.9979          
 Mcnemar's Test P-Value : NA              

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            1.0000   1.0000   0.9963   0.9953   0.9986
Specificity            1.0000   0.9995   0.9991   0.9994   1.0000
Pos Pred Value         1.0000   0.9980   0.9956   0.9969   1.0000
Neg Pred Value         1.0000   1.0000   0.9992   0.9991   0.9997
Prevalence             0.2845   0.1935   0.1744   0.1639   0.1838
Detection Rate         0.2845   0.1935   0.1737   0.1631   0.1835
Detection Prevalence   0.2845   0.1939   0.1745   0.1637   0.1835
Balanced Accuracy      1.0000   0.9998   0.9977   0.9974   0.9993
```
The accuracy is 99.83% which is pretty good prediction. In this case, we can use the random forest to predict the final prediction set with 20 cases. 

### Out of sample error
The out of sample error is the error rate you get on new data set. In my case, it's the error rate after running the predict() function on the 4 testing set. 

The expected out-of-sample error is 100%-99.83% = 0.17%.

### Prediction 
```{r}
prediction2=predict(modelfit1, finaltesting, type = "class")
prediction2

[1] B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 
```

### Conclusion

